Assessing exposure to Atlantic Basin tropical storms in United States counties

### 2016-06-30 writing/DraftExposurePaper/DraftExposurePaper.Rmd
Rainfall potential parameter

When assessing rainfall while a hurricane is over the ocean, @Cerveny2000 found that tropical cyclones occurring later in the hurricane season (e.g., October--January) were associated with significantly greater rates of precipitation (rain per day) than earlier in the season when evaluating storms in the North Pacific and Atlantic basins from 1979--1995. This obeservation is likely due to the high frequency of intense storms in the later months for the Pacific basin, seeing as the most intense storms often occur in September for the Atlantic basin [@Cerveny2000]. The rate of rainfall was found to be positively correlated to the wind speed (i.e., maximum surface wind speed) of a particular tropical cyclone, and therefore storm intensity [@Cerveny2000]. This is somewhat contradictory for the Atlantic basin, which often experiences its most intense storms early in the hurricane season (e.g., September) [@Cerveny2000].[JMF--All of the above results are based solely off observations over the ocean and therefore may not reflect storm characteristics once landfall has been made]. The relationship between precipitation and wind speed does not remain as strong once the hurricane has made landfall [@Jiang2008].

>>>>

When assessing rainfall while a hurricane is over the ocean, @Cerveny2000 found that tropical cyclones occurring later in the hurricane season (e.g., October--January) were associated with significantly greater rates of precipitation (rain per day) than earlier in the season when evaluating storms in the North Pacific and Atlantic basins from 1979--1995. This obeservation is likely due to the high frequency of intense storms in the later months for the Pacific basin, seeing as the most intense storms often occur in September for the Atlantic basin [@Cerveny2000]. The rate of rainfall was found to be positively correlated to the wind speed (i.e., maximum surface wind speed) of a particular tropical cyclone, and therefore storm intensity [@Cerveny2000]. This is somewhat contradictory for the Atlantic basin, which often experiences its most intense storms early in the hurricane season (e.g., September) [@Cerveny2000].[JMF--All of the above results are based solely off observations over the ocean and therefore may not reflect storm characteristics once landfall has been made]. The relationship between precipitation and wind speed does not remain as strong once the hurricane has made landfall [@Jiang2008]. In fact, though the intensity of a storm can be considerably reduced once landfall has been made, peak rainfalls overland tend to be greater than those over the ocean [@Jiang2008]. The area of peak rainfall overland seem to be close to the storm's center, about 50 km [@Jiang2008]. @Jiang2008 also found statistically significant relationships between decreasing rain rates from the center of the storm out to about 400 km, with no further significant relationship beyond this distance.

Researchers have attempted to predict the expected severity of rainfall for a given storm once it has made landfall using a rainfall potential parameter [@Griffith1978]. The rainfall potetnial parameter is developed by analyzing characteristics of the storm's evolution over the ocean, including: storm size, translational speed, and average rain rates derived from infrared satellite observations [@Griffith1978].
